---
title: "Getting data""
output: html_notebook
---

The idea is to build an animated time series of an ensmeble forecast using each member. Each member will be represented by
a dot and have a trajectory from time zero until the the time of interest. At the time of interest the dots are binned resuling 
in a "dots" plot as in the ggdist package. 

Let's start by getting some MEPS data that is quite uncertain - we could do that by taking 2m temperature from the most
recent MEPS forecast and computing the standard deviation at each lead time. 

```{r setup}
library(here)
library(harp)
library(dplyr)
library(purrr)
library(tidyr)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r get-intial-data}
t2m <- map(
  0:48,
  ~read_forecast(
    2020090800,
    2020090800,
    "meps",
    "T2m",
    lead_time        = .x,
    file_path        = here("data"),
    file_template    = "{fcst_model}/{YYYY}/{MM}/{DD}/meps_lagged_6_h_subset_2_5km_{YYYY}{MM}{DD}T{HH}Z.nc",
    file_format_opts = netcdf_opts(options_set = "met_norway_eps"), 
    return_data      = TRUE
  ) %>% 
    ens_stats(min = TRUE, max = TRUE) %>% 
    select(-contains("meps_mbr"))
)

t2m <- structure(
  set_names(list(bind_rows(map(t2m, 1))), "meps"),
  class = "harp_fcst"
)
```

It looks like there is some scope for some nice spread in the ensemble
```{r field-plot, message=FALSE, fig.align='center', warning=FALSE}
plot_field(t2m, "meps", ens_spread, lead_time = 42)
```

We can probably just get the data and interpolate to all stations and see what we get.
```{r interp-to-stations}
read_forecast(
  2020090800,
  2020090800,
  "MEPS_op",
  "T2m",
  lead_time        = seq(0, 48),
  file_path        = here("data/meps"),
  file_template    = "{YYYY}/{MM}/{DD}/meps_lagged_6_h_subset_2_5km_{YYYY}{MM}{DD}T{HH}Z.nc",
  file_format_opts = netcdf_opts(options_set = "met_norway_eps"), 
  transformation   = "interpolate",
  output_file_opts = sqlite_opts(path = here("data", "FCTABLE")),
  return_data      = TRUE, 
  show_progress    = TRUE 
)

read_forecast(
  2020090800,
  2020090800,
  "MEPS_op",
  "Pcp",
  lead_time        = seq(0, 48),
  file_path        = here("data/meps"),
  file_template    = "{YYYY}/{MM}/{DD}/meps_lagged_6_h_subset_2_5km_{YYYY}{MM}{DD}T{HH}Z.nc",
  file_format_opts = modify_at(netcdf_opts(options_set = "met_norway_eps"), "z_var", ~"height0"), 
  transformation   = "interpolate",
  output_file_opts = sqlite_opts(path = here("data", "FCTABLE")),
  return_data      = TRUE, 
  show_progress    = TRUE 
)
```

